---
title: "Weight Tracker"
output:
  html_document:
    df_print: paged
---

# Introduction

This will track and try to make sense of changes in weight over time. The data will have to be manually recorded on an Excel workbook.


# Packages etc

```{r}
pacman::p_load(
  openxlsx, tidyverse, plotly, lubridate
)

```
Data 


```{r}
data <- readxl::read_excel("weight_over_time.xlsx", sheet = "raw_data", col_types = c("date", "skip", "numeric", "skip"))

```

# Find AM and PM

```{r}
delta_data <- 
  data %>% 
    mutate(
      tod = ifelse(lubridate::am(date), "AM", "PM")
    ) %>% 
    mutate(
      day_alone = lubridate::as_date(date)
    ) %>% 
    select(day_alone, tod, weight) %>% 
    group_by(day_alone, tod) %>% 
    summarise( min_max_weight = case_when(
      tod[1] == "AM" ~ min(weight),
      tod[1] == "PM" ~ max(weight)
      )
    )

```

Get range of morning values 

```{r}
am_range <- 
  data %>% 
    filter(lubridate::am(date)) %>% 
    mutate(
        day_alone = lubridate::as_date(date)
    ) %>% 
      select(day_alone, weight) %>% 
      group_by(day_alone) %>% 
      summarise(
        min_am = min(weight),
        max_am = max(weight),
        n_meas = n()
      )

```

Now to visualise 


```{r, fig.height=6, fig.width = 8}

# regression model
mod <- lm(min_max_weight ~ day_alone + tod, data = delta_data)

# mod_gam <- gam::gam(min_max_weight ~ s(day_alone) + tod, data = ungroup(delta_data))

# get equation
# https://stackoverflow.com/questions/7549694/add-regression-line-equation-and-r2-on-graph

lm_eqn <- function(m){
  m_tidy <- broom::tidy(mod)
  m_glance <- broom::glance(mod)
  trend_point <- m_tidy %>% pluck(2, 2) %>% round(2)
  trend_se    <- m_tidy %>% pluck(3, 2) %>% `*`(2) %>% round(2)
  daily_point <- m_tidy %>% pluck(2, 3) %>% round(2) 
  daily_se    <- m_tidy %>% pluck(3, 3) %>% `*`(2) %>% round(2)
  adj_r_sq    <- m_glance$adj.r.squared[1]  %>% round(2)
  glue::glue("Daily fall of {trend_point} (\u00b1 {trend_se}) kg/ day\n
             Daily variation of {daily_point} (\u00b1 {daily_se}) kg\n
             Ratio: {round(abs(daily_point / trend_point), 2)}\n
             Adjusted R-squared: {adj_r_sq}
             ")
}

# First day of isoweeks 
first_isoweeks <- delta_data %>% 
  ungroup(day_alone) %>% 
  mutate(isoweek = lubridate::isoweek(day_alone)) %>% 
  mutate(ch_isoweek = isoweek != lag(isoweek)) %>% 
  filter(ch_isoweek) %>% 
  select(day_alone)

minimums <- 
  delta_data %>% 
  ungroup() %>% 
  filter(tod == "AM") %>% 
  mutate(min_so_far = cummin(min_max_weight)) %>% 
  mutate(change = min_so_far < lag(min_so_far)) %>% 
  filter(change | is.na(change))

delta_data %>% 
  modelr::add_predictions(., mod) %>% 
  ggplot(aes(x = day_alone)) + 
    geom_hline(yintercept = 72.7, linetype = "dashed", size = 1.5) + 
  geom_hline(yintercept = 69.0, linetype = "dashed", size = 1.0) +
  geom_hline(yintercept = 74.2, linetype = "solid", colour = "darkblue") +
  geom_vline(aes(xintercept = day_alone), data = first_isoweeks, linetype = "dashed", colour = "grey") + 
  geom_point(aes(y = min_max_weight, shape = tod, group = tod, colour = tod), alpha = 0.7) + 
  geom_point(aes(y = min_so_far, x = day_alone), data = minimums, shape = 0, size = 6, inherit.aes = FALSE) + 
  geom_line(aes(y = pred, group = tod, colour = tod)) +
  geom_line(aes(y = min_max_weight, x = day_alone), 
            colour = "grey", alpha = 0.5,
            inherit.aes = FALSE, 
            data = delta_data %>% 
              filter(tod == "AM")
              ) + 
  geom_linerange( # range of values within morning
    aes(x = day_alone, ymin = min_am, ymax = max_am), 
    inherit.aes = FALSE, data = am_range, 
    color = "#F8766D"
  ) +
  geom_point( # measures withing morning,
    aes(x = day_alone, y = weight), 
    inherit.aes = FALSE, 
    data = 
      data %>% 
      filter(lubridate::am(date)) %>% 
      mutate(
          day_alone = lubridate::as_date(date)
      ),
    alpha = 0.5, size = 0.25
  ) +
  scale_y_continuous(
    breaks = 65:80,
    sec.axis = sec_axis(trans = ~. / 6.35029, breaks = seq(10, 13, by = 0.1), name = "Weight in stones")                 
  ) + 
  scale_x_date(breaks = "1 week", 
               minor_breaks = "1 day",
               labels = scales::date_format("%d %b") 
               ) + 
  expand_limits(y = c(65, 80)) +
  labs(
    x = "Date",
    y = "Weight in kg",
    title = "Minimum morning and maximum evening weight over time",
    subtitle = "Blue line: BMI of 25. Thicker dashed line: 95% of initial weight; thinner: 90.25% of initial weight. Squares indicate new lowest weights",
    shape = "Time of day", colour = "Time of day"
  ) + 

  geom_text(x = as.Date("2020-09-01"), y = 77.5, label = lm_eqn(mod), colour = "black", hjust = 0) + 
  expand_limits(y = c(70, 80)) 


ggsave("weight_chart.png", height = 20, width = 30, units = "cm", dpi = 300)
  
```

And how about a projection over next 4 weeks?

```{r}
delta_data %>% 
  modelr::add_predictions(., mod) %>% 
  ggplot(aes(x = day_alone)) + 
  geom_hline(yintercept = 72.7, linetype = "dashed", size = 1.5) + 
  geom_hline(yintercept = 69.0, linetype = "dashed", size = 1.0) +
  geom_hline(yintercept = 74.2, linetype = "solid", colour = "darkblue") +
  geom_vline(aes(xintercept = day_alone), data = first_isoweeks, linetype = "dashed", colour = "grey") + 
  geom_point(aes(y = min_max_weight, shape = tod, group = tod, colour = tod), alpha = 0.7) + 
  geom_point(aes(y = min_so_far, x = day_alone), data = minimums, shape = 0, size = 6, inherit.aes = FALSE) + 
  geom_line(aes(y = pred, group = tod, colour = tod)) +
  geom_line(aes(y = min_max_weight, x = day_alone), 
            colour = "grey", alpha = 0.5,
            inherit.aes = FALSE, 
            data = delta_data %>% 
              filter(tod == "AM")
              ) + 
  geom_linerange( # range of values within morning
    aes(x = day_alone, ymin = min_am, ymax = max_am), 
    inherit.aes = FALSE, data = am_range, 
    color = "#F8766D"
  ) +
  geom_point( # measures withing morning,
    aes(x = day_alone, y = weight), 
    inherit.aes = FALSE, 
    data = 
      data %>% 
      filter(lubridate::am(date)) %>% 
      mutate(
          day_alone = lubridate::as_date(date)
      ),
    alpha = 0.5, size = 0.25
  ) +
  scale_y_continuous(
    breaks = 65:80,
    sec.axis = sec_axis(trans = ~. / 6.35029, breaks = seq(10, 13, by = 0.1), name = "Weight in stones")                 
  ) + 
  scale_x_date(breaks = "1 week", 
               minor_breaks = "1 day",
               labels = scales::date_format("%d %b") 
               ) + 
  expand_limits(y = c(65, 80)) +
  labs(
    x = "Date",
    y = "Weight in kg",
    title = "Minimum morning and maximum evening weight over time",
    subtitle = "Blue line: BMI of 25. Thicker dashed line: 95% of initial weight; thinner: 90.25% of initial weight. Squares indicate new lowest weights",
    shape = "Time of day", colour = "Time of day"
  ) + 

  expand_limits(y = c(70, 80))  + 
  geom_line(
    aes(x = day_alone, y = pred),
    inherit.aes = FALSE,
    data = modelr::add_predictions(
      data = tibble(
        day_alone = seq(
          lubridate::today(),
          lubridate::today() + weeks(6),
          by = '1 day'
        ),
        tod = "AM"
      ), 
      mod),
    linetype = "dashed"
  )


ggsave("weight_chart_projection.png", height = 20, width = 30, units = "cm", dpi = 300)
  
```

Appears to project 21 October. 


Let's try to calculate projected date to 69 kg against last date of observations


```{r}
project_to_weight <- function(dta, last_date, target_weight = 69, window = seq(ymd("2020-08-01"), ymd("2020-12-31"), by = '1 day')){
  mod <- dta %>% 
    filter(day_alone <= last_date) %>% 
    lm(min_max_weight ~ day_alone + tod, data = .)

  fake_data <- tibble(
    day_alone = window, tod = "AM"
  ) 
  
  preds <- modelr::add_predictions(fake_data, mod)
  
  pred_target_date <- preds %>% 
    filter(pred < target_weight) %>% 
    pull(day_alone) %>% 
    pluck(1) %>% 
    as.character()
  
  return(pred_target_date)
}


projections_by_date <- 
  tibble(
    last_date = seq(
      dmy("1-9-2020"), today(), by = '1 day'
    )
  ) %>% 
  mutate(projection_date = map_chr(last_date, project_to_weight, dta = delta_data)) %>% 
  mutate(projection_date  = ymd(projection_date))




```

Visualise

```{r}
projections_by_date %>% 
  ggplot(aes(last_date, projection_date)) +
  geom_point() + 
  geom_line() + 
  scale_x_date(breaks = "1 week", 
               minor_breaks = "1 day",
               labels = scales::date_format("%d %b") 
  ) + 
  scale_y_date(breaks = "1 week", 
               minor_breaks = "1 day",
               labels = scales::date_format("%d %b") 
  ) +
  labs(
    x = "Last date included in projection",
    y = "First date on which 69kg weight first achieved"
  )

```

And linear regression

```{r}
mod <- lm(min_max_weight ~ day_alone + tod, data = delta_data)

```

```{r}
summary(mod)

```


So, on the basis of this there's a statistically significant downward trend of about 0.09 kg / day, but this is dwarfed by daily variation which is around 1.3 kg / day, i.e. 10x its magnitude


```{r}
mod2 <- lm(min_max_weight ~ day_alone, data = delta_data)

summary(mod2)

anova(mod, mod2)

```
Any weekend effect? 

```{r}
mod3 <- lm(min_max_weight ~ day_alone + is_weekend, 
           data = delta_data %>% 
             mutate(is_weekend = case_when(
               lubridate::wday(day_alone) %in% c(1, 7) ~ TRUE,
               TRUE ~ FALSE
             ))
             )

summary(mod3)

anova(mod2, mod3)

```
No convincing evidence ( p < 0.01 etc) of a systemic weekend effect, though indications of slightly higher weekend weights. (+ 0.3 kg) 


# Weekly summaries 


```{r}
mdls_by_week <- 
  delta_data %>% 
    mutate(
      week = lubridate::isoweek(day_alone)
    ) %>% 
    group_by(week) %>% 
    nest() %>% 
    mutate(n = map_dbl(data, nrow)) %>%  
    filter(n > 3) %>%  
    mutate(mdl = map(.x = data, ~lm(min_max_weight ~ day_alone + tod, data = .))) %>% 
    mutate(mdl_tidy = map(mdl, broom::tidy)) %>% 
    mutate(mdl_sum = map(mdl, broom::glance))


mdls_by_week %>% 
  select(week, mdl_tidy) %>% 
  unnest(mdl_tidy) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(
    estimate = ifelse(term == "day_alone", estimate * 7, estimate),
    std_error = ifelse(term == "day_alone", std.error * 7, std.error)
  ) %>% 
  mutate(lower = estimate - 2 * std.error,
         upper = estimate + 2 * std.error) %>% 
  select(week, term, estimate, lower, upper) %>% 
  left_join(
    mdls_by_week %>% 
      select(week, mdl_sum) %>% 
      unnest(mdl_sum) %>% 
      ungroup() %>%  
      select(week, adj.r.squared)
  ) %>% 
  filter(week > 31) %>% 
  mutate(Regimen = case_when(
    week <  36 ~ "Time-restricted feeding",
    week >= 36 ~ "Vegetarian Keto",
    TRUE       ~ NA_character_
    ) 
  ) %>% 
  mutate(term = case_when(
    term == "day_alone" ~ "Change in week",
    term == "todPM"       ~ "Within-day variation",
    TRUE                    ~ NA_character_
  )) %>% 
  ggplot(aes(x = week, colour = Regimen)) + 
  geom_point(aes(y = estimate, size = adj.r.squared, shape = Regimen)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) + 
  facet_wrap(~term)  +
  geom_hline(yintercept = 0) + 
  labs(
    x = "Week of year",
    y = "Estimate in kg",
    title = "Coefficients on trend and within-day variation by week",
    subtitle = "Error bars: Two standard errors. Point size proportional to model fit"
  ) + 
  guides(size = "none") + 
  scale_y_continuous(breaks = seq(-2.5, 2.5, by = 0.5), minor_breaks = seq(-2.5, 2.5, by = 0.1))

ggsave("coeffs_by_week.png", height = 20, width = 30, units = "cm", dpi = 300)








```

Lagged fortnight?

```{r}

filter_fortnight <- function(x, D = delta_data){
  D %>% 
    mutate(week = lubridate::isoweek(day_alone)) %>% 
    filter(week %in% c(x, x - 1))
}

mdls_by_fortnight <- 
  tibble(
    week = delta_data %>% # actually week ending 2 week period
      mutate(week = lubridate::isoweek(day_alone)) %>% 
      pull(week) %>% unique()         
  ) %>% 
  mutate(
    data = map(week, filter_fortnight)
  ) %>% 
  mutate(n = map_dbl(data, nrow)) %>%  
  filter(n > 3) %>%  
  mutate(mdl = map(.x = data, ~lm(min_max_weight ~ day_alone + tod, data = .))) %>% 
    mutate(mdl_tidy = map(mdl, broom::tidy)) %>% 
    mutate(mdl_sum = map(mdl, broom::glance))


mdls_by_fortnight %>% 
  select(week, mdl_tidy) %>% 
  unnest(mdl_tidy) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(
    estimate = ifelse(term == "day_alone", estimate * 7, estimate),
    std_error = ifelse(term == "day_alone", std.error * 7, std.error)
  ) %>% 
  mutate(lower = estimate - 2 * std.error,
         upper = estimate + 2 * std.error) %>% 
  select(week, term, estimate, lower, upper) %>% 
  left_join(
    mdls_by_week %>% 
      select(week, mdl_sum) %>% 
      unnest(mdl_sum) %>% 
      ungroup() %>%  
      select(week, adj.r.squared)
  ) %>% 
  filter(week > 31) %>% 
  mutate(Regimen = case_when(
    week <  36 ~ "1. Time-restricted feeding",
    week == 36 ~ "2. Mixed",
    week >  36 ~ "3. Vegetarian Keto",
    TRUE       ~ NA_character_
    ) 
  ) %>% 
  mutate(term = case_when(
    term == "day_alone" ~ "Change in week",
    term == "todPM"       ~ "Within-day variation",
    TRUE                    ~ NA_character_
  )) %>% 
  ggplot(aes(x = week, colour = Regimen)) + 
  geom_point(aes(y = estimate, size = adj.r.squared, shape = Regimen)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) + 
  facet_wrap(~term)  +
  geom_hline(yintercept = 0) + 
  labs(
    x = "Week of year (Last week)",
    y = "Estimate in kg",
    title = "Coefficients on trend and within-day variation by fortnights",
    subtitle = "Error bars: Two standard errors. Point size proportional to model fit"
  ) + 
  guides(size = "none") + 
  scale_y_continuous(breaks = seq(-2.5, 2.5, by = 0.5), minor_breaks = seq(-2.5, 2.5, by = 0.1))

ggsave("coeffs_by_fortnight.png", height = 20, width = 30, units = "cm", dpi = 300)





```


# Change in measurements within day 

```{r}

data %>% 
  mutate(is_am = am(date)) %>% 
  filter(is_am) %>% 
  mutate(
    date_alone = lubridate::floor_date(date, "day"),
    hod = lubridate::hour(date) + lubridate::minute(date) / 60
  ) %>% 
  mutate(days_since_today = round(interval(date_alone, today()) %/% days(1))) %>% 
  ggplot(
    aes(
      x = hod, y = weight, group = date_alone, alpha = scales::rescale(days_since_today, to = c(1, 0))
  )) + 
  geom_point() + 
  geom_line() + 
  guides(alpha = "none")
  

```